\sys: Light-touch Data-Parallel Shell Processing

### 2020-10-08 paper/draft.tex
Update 6.2

We now evaluate \sys on an existing set of \unix pipelines found in the wild.

\heading{Programs}
In a recent celebration of \unix's 50-year legacy, Bell Labs created 37 challenges~\cite{unix50} solvable by \unix pipelines.
The problems were designed to highlight \unix's modular philosophy~\cite{mcilroy1978unix}.
We found unofficial solutions to all-but-three problems on GitHub~\cite{unix50sol}, expressed as pipelines with 2--12 stages (avg.: 5.58).
They make extensive use of standard commands under a variety of flags, and appear to be written by non-experts (contrary to \S\ref{ours}, they often use sub-optimal or non-\unix-y constructs).
% ---stressing a patterns that we, \sys's authors, did not
% necessarily anticipate \kk{I would change that to ``stressing a
%   variety of patterns'' since it doesn't really matter if we
%   anticipated them (especially since we can address any issues and
%   then re-run them).} . 
\sys executes each pipeline as-is, without any modification.
  % if an obvious fix improves performance, we report the speedup of both versions. %  original and modified. %  pipeline.

@@

\heading{Results}
\sys's \ttt{width} is set to 16 and pipelines are given 10GB inputs.
Fig.~\ref{fig:unix50-individual} shows the speedup (left) over the sequential runtime (right) for 31 pipelines.
We do not include three pipelines that take less than 0.1 seconds to execute because they only output the first line of their input using \ttt{head}.
We use the pipeline index to refer to them in both plots and discussion.
Average speedup is \todo{$6.02\times$}, median is $6.48\times$, and weighted average (with the absolute times as weights) is $5.75\times$.
 % the \todo{geometric mean is $1.08\times$},

@@

\heading{Discussion}
Most pipelines see significant acceleration, except 25-30
  % \todo{13, 24, 25, 26, 29, 30} 
  that see no speedup. 
  % and \todo{2, 19, 31} that see a slowdown.
These pipelines often contain general commands that \sys cannot parallelize without risking breakage---\eg \ttt{awk} and \ttt{sed} \ttt{-d}. % and \ttt{tr} \ttt{-d}.
% Note however, that some of the pipelines contain more general commands that cannot be parallelized since they were not written with that goal in mind;
%   a common example is the use of \ttt{awk} for reordering or removing columns of the input.
%% Some of them use general interpreters when they could be using a specialized equivalent tailored to the task at hand, the typical example being \ttt{awk} vs \ttt{cut}.
A \unix expert would notice that some of these uses could be replaced with \unix-specific commands---\eg replacing \ttt{awk "{print \$2, \$0}" | sort -nr} just used to sort on the second field with a single \ttt{sort -nr -k 2} in the 26th pipeline.
  % \todo{13th} 
The limited expressiveness of these commands can be exploited by \sys---in this specific case, achieving 8.1$\times$ speedup (\vs the original 1.01$\times$).
%% , attempting only such obvious (to us, experts) replacements, \sys extracts further parallelism---\eg 13 sees a 8.1$\times$ speedup.
% ---a common pattern we witnessed~\sx{macro1}.

@@

For the rest, \sys's speedup is capped due to a combination of reasons:
  (i) they contain pure commands that are parallelizable but don't scale linearly, such as \ttt{sort} 
  (\# 5, 6, 7, 8, 9, 19, 20, 21, 23, 24), 
  % 24, 9, 7, 6, 8, 23, 20, 19, 21, 5
  % \todo{(0, 1, 3, 15, 16, 18, 20, 27, 28, 33)},
  (ii) they are deep pipelines that already exploit task-based parallelism
  (\# 4, 10, 11, 13, 15, 17, 19, 21, 22),
  % 17, 4, 13, 15, 10, 22, 11, 19, 21,
  % \todo{(7, 8, 9, 10, 11, 12, 14, 27, 28,)} 
or 
  (iii) they are not CPU-intensive, resulting in pronounced IO and constant costs 
  (\# 3, 4, 11, 12, 14, 16, 17, 18, 22).
  % 14, 3, 16, 17, 4, 22, 11, 12, 18
  % \todo{(4, 5, 6, 7, 8, 12, 14, 22, 23)}.

>>>>

We now turn to a set of \unix pipelines found out in the wild.

\heading{Programs}
In a recent celebration of \unix's 50-year legacy, Bell Labs created 37 challenges~\cite{unix50} solvable by \unix pipelines.
The problems were designed to highlight \unix's modular philosophy~\cite{mcilroy1978unix}.
We found unofficial solutions to all-but-three problems on GitHub~\cite{unix50sol}, expressed as pipelines with 2--12 stages (avg.: 5.58).
They make extensive use of standard commands under a variety of flags, and appear to be written by non-experts (contrary to \S\ref{ours}, they often use sub-optimal or non-\unix-y constructs).
% ---stressing a patterns that we, \sys's authors, did not
% necessarily anticipate \kk{I would change that to ``stressing a
%   variety of patterns'' since it doesn't really matter if we
%   anticipated them (especially since we can address any issues and
%   then re-run them).} . 
\sys executes each pipeline as-is, without any modification.\footnote{
  \nv{Do we need this?}
}
  % if an obvious fix improves performance, we report the speedup of both versions. %  original and modified. %  pipeline.

@@

\heading{Results}
Fig.~\ref{fig:unix50-individual} shows the speedup (left) over the sequential runtime (right) for 31 pipelines, with \ttt{width=}16 and 10GB inputs.
It does not include 3 pipelines that execute in under $0.1s$ by using \ttt{head} to output the first line of input.
The script index refers to the challenge level, in both the plots and the discussion.
Average speedup is \todo{$6.02\times$}, % median is $6.48\times$, 
  and weighted average (with the absolute times as weights) is $5.75\times$.
 % the \todo{geometric mean is $1.08\times$},

@@

\heading{Discussion}
Most pipelines see significant speedup, except \#25-30
  % \todo{13, 24, 25, 26, 29, 30} 
  that see no speedup because they contain general commands that \sys cannot parallelize without risking breakage---\eg \ttt{awk} and \ttt{sed} \ttt{-d}. % and \ttt{tr} \ttt{-d}.
  % and \todo{2, 19, 31} that see a slowdown.
% Note however, that some of the pipelines contain more general commands that cannot be parallelized since they were not written with that goal in mind;
%   a common example is the use of \ttt{awk} for reordering or removing columns of the input.
%% Some of them use general interpreters when they could be using a specialized equivalent tailored to the task at hand, the typical example being \ttt{awk} vs \ttt{cut}.
A \unix expert would notice that some of them can be replaced with \unix-specific commands---\eg \ttt{awk "{print \$2, \$0}" | sort -nr}, used to sort on the second field can be replaced with a single \ttt{sort -nr -k 2} (\#26).
  % \todo{13th} 
The limited expressiveness of these commands can be exploited by \sys---in this specific case, achieving 8.1$\times$ speedup (\vs the original 1.01$\times$).
%% , attempting only such obvious (to us, experts) replacements, \sys extracts further parallelism---\eg 13 sees a 8.1$\times$ speedup.
% ---a common pattern we witnessed~\sx{macro1}.

@@

For the rest, \sys's speedup is capped due to a combination of reasons:
  (i) they contain pure commands that are parallelizable but don't scale linearly, such as \ttt{sort} 
  (\#5, 6, 7, 8, 9, 19, 20, 21, 23, 24), 
  % 24, 9, 7, 6, 8, 23, 20, 19, 21, 5
  % \todo{(0, 1, 3, 15, 16, 18, 20, 27, 28, 33)},
  (ii) they are deep pipelines that already exploit task parallelism
  (\#4, 10, 11, 13, 15, 17, 19, 21, 22),
  % 17, 4, 13, 15, 10, 22, 11, 19, 21,
  % \todo{(7, 8, 9, 10, 11, 12, 14, 27, 28,)} 
or 
  (iii) they are not CPU-intensive, resulting in pronounced I/O and constant costs 
  (\#3, 4, 11, 12, 14, 16, 17, 18, 22).
  % 14, 3, 16, 17, 4, 22, 11, 12, 18
  % \todo{(4, 5, 6, 7, 8, 12, 14, 22, 23)}.
